---
title: "Getting Data"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkgs, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(tidycensus)
# I'm watching Love, Actually while I make this. :) Happy holidays! 
```

##  Types of Data 

For this project we will need:

- Census Data: to count populations in areas 
- Shapefiles (also can be from Census): to compute the geometric measures 
- State-level election data: precinct vote counts, also shapefiles 


### Census Data 

The [tidycensus](https://walker-data.com/tidycensus/) package is an interface to the Census API that gets data in a [tidy](https://vita.had.co.nz/papers/tidy-data.pdf) way. The intro vignette to the package is [here](https://walker-data.com/tidycensus/articles/basic-usage.html). 

1. Get an API key from [here](https://api.census.gov/data/key_signup.html)
2. Install a couple packages:
```r
install.packages(c("tidyverse", "tidycensus", "usethis")) 
```
3. Add the API key emailed to you to your `.Renviron` file. To open this file, run:
```r
usethis::edit_r_environ()
```
On a new line in the .Renviron file, put: 
```r
CENSUS_API_KEY="yourkeyhere"
```
4. Restart R. (Session -> Restart R)
5. Run the following: 
```r
library(tidyverse)
library(tidycensus)

census_api_key(Sys.getenv("CENSUS_API_KEY"), install = TRUE)
```
6. Now you can use `tidycensus`! 

#### Variables

The Census records a whole lotta variables, and `tidycensus` has a way to search for them: 
```r
v19 <- load_variables(2019, "acs5", cache = TRUE)
View(v19)
```
The 2019 ACS data was just released! 

Use this call to get total population in the state legislative districts of Virginia in 2019 5-year ACS data. 


```{r}
pop_var <- "B01003_001"
pop_st_leg <- get_acs(geography = "state legislative district (upper chamber)",
                       variables = c(total_pop = pop_var),
                       state="VA", 
                       year = 2019)

head(pop_st_leg)
```

### Shapefiles 

The [`tigris`](https://github.com/walkerke/tigris) package will allow us to directly query the TIGER/Line shapefiles from the Census Bureau.

1. Install some packages: 
```r
install.packages(c('tigris', 'sf'))
```
2. 
```{r}
library(tigris)
library(sf)

va_st_leg <- state_legislative_districts(state = "VA", 
                            house = "upper", 
                            year = 2019)
va_st_leg
```

### State-level election data

The website [openprecincts.org](https://openprecincts.org/va/) has state-level precinct data. The latest election data for Virginia is from 2018. The data for the Senate election is in the [dat](dat/) folder. More results for 2018 can be downloaded [here](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/UBKYRU&version=4.0). 

1. Read in the shape file: 
```{r, message=FALSE}
va_2018_sen <- st_read("dat/va_2018/va_2018_senate.shp") #sf package 
va_2018_sen
```

## Visualizing Data 

The combination of [`sf`](https://r-spatial.github.io/sf/) and [`ggplot2`](https://ggplot2.tidyverse.org/) is pretty magical here. 

```{r}
ggplot(data = va_st_leg) + 
  geom_sf(size = .2) + 
  ggtitle("Virginia State Senate Districts") + 
  theme_void()
```

### Combining data 

Pull in the population data from `po_st_leg`: 

```{r}
va_st_leg2 <- va_st_leg %>% 
  left_join(pop_st_leg, by = c("GEOID" = "GEOID"))

ggplot(data = va_st_leg2) + 
  geom_sf(size = .2, aes(fill = estimate)) +
  scale_fill_distiller(name = "Population", palette = "PuBu", direction = 1) + 
  ggtitle("Virginia State Senate Districts", subtitle = "Population Estimate, ACS 2019 5-yr") + 
  theme_void()
```

